---
title: "assignment"
author: "Jesslyn Jesslyn"
date: "2025-03-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
suppressPackageStartupMessages({
  library(rtracklayer)
  library(epiwraps)
})
```

## Using peaks from ENCODE, identify bivalent domains (H3K27me3 + H3K4me3) from (mouse or human) embryonic stem cells (ESC)
Using mouse ESC:

```{r}
download.file("https://www.encodeproject.org/files/ENCFF008XKX/@@download/ENCFF008XKX.bed.gz", "h3k27me3_esc.bed.gz")
h3k27me3_esc <- rtracklayer::import("h3k27me3_esc.bed.gz", format="narrowPeak")

download.file("https://www.encodeproject.org/files/ENCFF974BMC/@@download/ENCFF974BMC.bed.gz", "h3k4me3_esc.bed.gz")
h3k4me3_esc <- rtracklayer::import("h3k4me3_esc.bed.gz", format="narrowPeak")
```

```{r}
## find bivalent domains
esc_overlap <- findOverlapPairs(h3k27me3_esc, h3k4me3_esc)
head(esc_overlap)

esc_bivalent <- punion(esc_overlap)
head(ranges(esc_bivalent))
```

*Question: what happens to those regions upon differentiation?*

-   Choose *one* differentiated cell type (e.g. erythroblasts, fibroblast, B cells, etc.)

Myoblast cells: C2C12 cell line

-   Download the H3K27me3 and H3K4me3 peaks from this cell type
```{r}
download.file("https://www.encodeproject.org/files/ENCFF071FEL/@@download/ENCFF071FEL.bed.gz", "h3k27me3_diff.bed.gz")
h3k27me3_diff <- rtracklayer::import("h3k27me3_diff.bed.gz", format="narrowPeak")

download.file("https://www.encodeproject.org/files/ENCFF635VQE/@@download/ENCFF635VQE.bed.gz", "h3k4me3_diff.bed.gz")
h3k4me3_diff <- rtracklayer::import("h3k4me3_diff.bed.gz", format="narrowPeak")
```

-   How many of the mESC bivalent domains are, in this differentiated cell type, overlapping either mark or their combination (in this differentiated cell type)?

1. 1972 of the mESC bivalent domains overlap with H3K27me3 peaks of the differentiated cell line.
2. 1133 of the mESC bivalent domains overlap with H3K4me3 peaks of the differentiated cell line.
3. 224 of the mESC bivalent domains overlap with both H3K27me3 and H3K4me3 peaks of the differentiated cell line.

```{r}
peaks <- list(h3k27me3_diff, h3k4me3_diff, esc_bivalent)
names(peaks) <- c("H3K27me3_differentiated", "H3K4me3_differentiated", "mESC_bivalent")
regionUpset(peaks, reference = peaks$mESC_bivalent, nsets=length(peaks))
```

(You can use either mouse or human data, just make sure you do everything with the same species!)
